### 2016-04-14 chapter3.tex

\section{炉石传说人工智能}
\label{section:HearthstoneAI}
\subsection{炉石传说基本规则}
\label{section:HearthstoneRule}